---
<<<<<<< HEAD
title: "Personal Activity Report"
author: "Hien Nguyen"
date: "Sunday, July 12, 2015"
output: html_document
---
Data
*The data for this assignment can be downloaded from https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip*

*Dataset: Activity monitoring data [52K]*
*The variables included in this dataset are:*

*steps: Number of steps taking in a 5-minute interval (missing values are coded       as NA)*

*date: The date on which the measurement was taken in YYYY-MM-DD format*

*interval: Identifier for the 5-minute interval in which measurement was taken*

*The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.*

Load the data into an activityOrig dataset
```{r, echo=TRUE}
activityOrig <- read.csv("activity.csv")
```
Format the date
```{r, echo=TRUE}
activityOrig$date <- as.Date(activityOrig$date, "%Y-%m-%d")
```
<font color="blue"><b>QUESTION 1:   What is mean total number of steps taken per day?</b></font>
<br>Create a new dataset from activityOrig without NAs
```{r, echo=TRUE}
activity <- na.omit(activityOrig)
```
Calculate the total steps per day and put in the column "total"
```{r, echo=TRUE}
require("plyr")
sumByDate <- ddply(activity, .(date), summarise, total=sum(steps))
```
Create a histogram of the total steps and run a summary to get mean/median
```{r, echo=TRUE}
mygraph <- hist(as.numeric(sumByDate$total), col="red",main= "Histogram of Total Steps", xlab="Number of Steps", ylab="Frequency")
summary(sumByDate)
```
<font color="blue">ANSWER 1</font>
**Median :10765**
**Mean   :10766**
<p><font color="blue"><b>QUESTION 2: What is the average daily activity pattern?</b></font>
<br>To answer this question, let's create a subset with just 2 columns: steps and intervals
```{r, echo=TRUE}
stepInterval <- subset(activity, select = c(steps, interval))
```
From this subset, we are calculating the average of steps for each interval
```{r, echo=TRUE}
averageForEachInterval <- ddply(stepInterval, .(interval), summarise, avg=mean(steps))
```
plot the average steps per interval
```{r, echo=TRUE}
mygraph2 <- with(averageForEachInterval, plot(interval, avg, main= "Average steps per interval", ylab="Average", xlab="Intervals", type="l"))
```
<P>let's find out what's the max average
```{r, echo=TRUE}
summary(averageForEachInterval$avg)
```
The max is 206.170.  This was a round up number
<br>To find which interval corresponds to this max value
```{r, echo=TRUE}
averageForEachInterval[as.numeric(averageForEachInterval$avg) > 206,]
```
<font color="blue">ANSWER: Interval 835 has the highest average of steps of 206.170</font>

<p><font color="blue"><b>QUESTION 3: Imputing missing values</b></font>
<br>1.  Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r, echo=TRUE}
activityNA <- subset(activityOrig, is.na(steps))
nrow(activityNA)
```
<font color="blue">Answer:  2304 rows with NA values.</font>

<br>2.  Filling in all of the missing values in the dataset with the mean for that 5-minute interval
<br>a.  Add the average column
```{r, echo=TRUE}
activityImput <- merge(activityNA, averageForEachInterval,by="interval")
```
<br>b.  Remove the NA column
```{r, echo=TRUE}
activityImput <- subset(activityImput, select = -c(2) )
```
<br>c.  Rename the average column to "steps"
```{r, echo=TRUE}
colnames(activityImput)[3] <- "steps"
```
<br>d.  Reorder the columns
```{r, echo=TRUE}
activityImput <- subset(activityImput, select = c(3,2,1))
```
<br>3.  Create a new dataset that is equal to the original dataset but with the missing data filled in by combining activityImput and activity.
```{r, echo=TRUE}
activityNew <- rbind(activity, activityImput)
```
<br> 4. Make a histogram of the total number of steps taken each day
<br>Calculate the total steps per day and put in the column "total"
```{r, echo=TRUE}
sumByDate2 <- ddply(activityNew, .(date), summarise, total=sum(steps))
```
Create a histogram of the total steps and run a summary to get mean/median
```{r, echo=TRUE}
mygraph2 <- hist(as.numeric(sumByDate2$total), col="green",main= "Histogram of Total Steps", xlab="Number of Steps", ylab="Frequency")
summary(sumByDate2)
```
<font color="blue">ANSWER 3</font>
**Median :10766**
**Mean   :10766**
<br> Do these values differ from the estimates from the first part of the assignment? 
<br> The mean stays the same.  The median is up by 1.
<br> What is the impact of imputing missing data on the estimates of the total daily number of steps?
<br> very slightly
<font color="blue"><b>QUESTION 4:  Are there differences in activity patterns between weekdays and weekends?</b></font>
<br> first, add a column to store the day of the week
```{r, echo=TRUE}
activityNew$weekday <- weekdays(as.Date(activityNew$date))
```
<br> then, put the Saturday and Sunday records in one table, mark them "weekend"
```{r, echo=TRUE}
activityWeekend <- activityNew[activityNew$weekday %in% c("Saturday", "Sunday"),]
activityWeekend$weekday <- "weekend"
```
<br> then, put the rest of the records in one table, mark them "weekday"
```{r, echo=TRUE}
activityWeekday <- activityNew[activityNew$weekday %in% c("Monday", "Tuesday","Wednesday", "Thursday","Friday"),]
activityWeekday$weekday <- "weekday"
```
<br> calculate the average for activityWeekend and activityWeekday
```{r, echo=TRUE}
avgPerIntervalWeekend <- ddply(activityWeekend, c("interval", "weekday"), summarise, avg=mean(steps))
avgPerIntervalWeekday <- ddply(activityWeekday, c("interval", "weekday"), summarise, avg=mean(steps))
avgPerIntervalAll <- rbind(avgPerIntervalWeekend, avgPerIntervalWeekday)
mygraph3 <- with(avgPerIntervalWeekend, plot(interval, avg, main= "Average steps per interval (weekend)", ylab="Average", xlab="Intervals", type="l"))
mygraph4 <- with(avgPerIntervalWeekday, plot(interval, avg, main= "Average steps per interval (weekday)", ylab="Average", xlab="Intervals", type="l"))
#mygraph5 <- with(avgPerIntervalAll, plot(interval, avg, main= "Average steps per interval (weekday)", ylab="Average", #xlab="Intervals", type="l", col=weekday))
head(avgPerIntervalAll)
```
graph
```{r, echo=TRUE}
require(ggplot2)

#mygraph5 <-- xyplot(interval~avg|weekday, data=avgPerIntervalAll, type = "l")
ggplot(avgPerIntervalAll, aes(interval, avg, colour = weekday)) + geom_line() + facet_grid(.~weekday) + facet_wrap(~weekday, nrow=2) +  xlab("Interval") + ylab("Average Number of Steps")
```
=======
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data



## What is mean total number of steps taken per day?



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
>>>>>>> 80edf39c3bb508fee88e3394542f967dd3fd3270
